Sistema Online de Filtragem em um Ambiente com Alta Taxa de Eventos

### 2006-11-05 published.tex
Theme release candidate 1
git-svn-id: file:///Users/andre/svn/andre_phd/thesis/trunk@119 b2e9302d-b4d2-0310-80c4-8586bce995ec
Converted all text to UTF-8
git-svn-id: file:///Users/andre/svn/andre_phd/thesis/trunk@327 b2e9302d-b4d2-0310-80c4-8586bce995ec
Update with provisional (but incomplete) publication list
git-svn-id: file:///Users/andre/svn/andre_phd/thesis/trunk@394 b2e9302d-b4d2-0310-80c4-8586bce995ec

\chapter{Publicações e Projetos}
\label{chap:published}

Durante a execução deste estudo, alguns trabalhos foram publicados na forma de
artigos e notas técnicas. Ademais, a participação da UFRJ, através da estadia
do doutorando no CERN, foi concretizada através de projetos realizados em
parceria direta. Neste capítulo, descreve-se cada um dos itens relevantes a
este trabalho. O período coberto inicia-se em abril de 2001 e, estende-se até
o presente momento. As citações estão organizadas pela data de publicação e,
no caso de projetos, pela data de início.

\section{Trabalhos Publicados}

@@

\paragraph{Architecture of the ATLAS High Level Trigger Event Selecti
on Software} Este trabalho foi apresentado no congresso \sectioneng{Computing
for High-Energy Physics' 2003}, em \sectioneng{La Jolla}, Califórnia,
EUA. Referência \cite{aa:chep-2003-3}

@@

\paragraph{Outubro de 2003: \sectioneng{Architecture of the ATLAS Online
Physics-Selection Software at LHC}}. Este artigo foi apresentado no
\sectioneng{8th ICATPP Conference Proceedings} em Villa Erba, Como, Itália. 
Referência \cite{aa:como-2003}.

\paragraph{Junho de 2004: \sectioneng{The base-line DataFlow system of the ATLAS
Trigger and DAQ}} A ser publicado na revista \sectioneng{IEEE Transactions on
Nuclear Science}, volume 51, número 3. A revista será publicada em junho de
2004 e, portanto, o número das páginas ainda não está disponível. Referência
\cite{aa:tns-2004-3}.

\section{Projetos}

Durante a estadia no CERN, embora a partipação tenha se dado no escopo do
projeto do sistema de filtragem, alguns tópicos específicos tiveram especial
atenção. Nesta seção destacam-se as principais contribuições durante o período
do doutorando no exterior.

Os projetos continuam ativos embora, em alguns casos, tenham atingido
maturidade e requeiram somente esforço de manutenção e correção de
problemas. Outros ainda estão em fase de desenvolvimento. Distinguem-se, no
decorrer do texto, projetos que chegaram à fase de estabilidade operacional
para o Sistema de Filtragem, daqueles que ainda estão sendo produzidos.

\paragraph{A Unidade de Processamento do LVL2}

Este projeto foi desenvolvido dentro da comunidade de Fluxo de Dados. Como
explanado anteriormente, a Unidade de Processamento do LVL2 ou L2PU é parte
integrante do sistema de aquisição e hospeda os algoritmos de filtragem do
ambiente Athena, enquanto executados \eng{online}.

Este projeto teve início em janeiro de 2002 e foi considerado maduro em meados
de 2003. Não há mais desenvolvimentos nesta área. O projeto é apenas mantido e
incrementado para a correção de erros e adição de pequenos ajustes.

\paragraph{Pseudo-ROS}

Este projeto foi desenvolvido dentro da comunidade de Fluxo de Dados. O
Pseudo-ROS recebe os dados dos eventos processados por uma L2PU. Grande parte
do código deste processador foi desenvolvida em parceria com a UFRJ.

Este projeto teve início em meados de 2002 e foi considerado maduro em meados
de 2003. Não há mais desenvolvimentos nesta área. O projeto é apenas mantido e
incrementado para a correção de erros e adição de pequenos ajustes.

\paragraph{Biblioteca de Formatação de Eventos}

Este projeto foi desenvolvido dentro da comunidade de Fluxo de Dados. Esta
biblioteca pode ler e escrever eventos gerados pelo detetor ATLAS e é usada
por toda a comunidade do experimento (\eng{online} e \eng{offline}) em
múltiplas aplicações. Este produto foi totalmente desenvolvido pela
Universidade e encontra-se em fase de manutenção apenas. Ele foi considerado
maduro em meados de 2003.

\paragraph{Relatório de Erros}

Este projeto está sendo desenvolvido dentro da comunidade de Fluxo de
Dados. Esta biblioteca proverá, quando acabada, uma interface amigável para
que as aplicações do Fluxo de Dados (L2PU, L2SV, PROS, ROS, etc) possam
relatar erros ou avisos decorrentes de situações inusitadas. Este sistema
estará intimamente ligado ao sistema de controle do experimento, que contará
ultimamente com um programa controlador, suficientemente inteligente, que
aplicará ações corretivas ao Sistema de Filtragem, com intervenção mínima do
operador do \eng{run}.

A biblioteca de relatório de erros começou a ser desenvolvida em novembro de
2003 e tem data prevista de término para outubro de 2004.

\paragraph{Biblioteca de Pré-processamento de Dados de Calorímetros}

Este projeto implementou uma biblioteca rápida de pré-processamento de dados
para os Calorímetros do ATLAS. Este projeto teve o único objetivo de provar
que era possível vencer os enormes tempos de processamento requeridos pelo
ambiente Athena. Os resultados tiveram um impacto bastante positivo na
comunidade \eng{Offline}, tendo a maior parte do código sido transportada para
aquele conjunto de bibliotecas. Este projeto conseguiu reduzir, em quase 50
vezes, o tempo de processamento para objetos calorimétricos. Embora continue
presente no Sistema de Filtragem, o projeto em si está inativo.

\paragraph{Athena Multi-threaded}

Este projeto foi encaminhado com a participação da Universidade. O projeto
consiste do transporte das rotinas do ambiente \eng{Offline}, para que o
Athena possa ser executado dentro do Sistema de Filtragem. Este projeto rendeu
muitos frutos e, também, uma publicação em revista (\eng{Studies for a common
Event Selection Software: from LVL2 to Offline reconstruction}) especializada,
como mencionado na seção anterior.

Este projeto foi iniciado em meados de 2002 e continua ativo. Não é possível
prever o tempo que demorará até que amadureça e possa ser usado em escala de
produção.

>>>>

\chapter{Publicações}
\label{chap:published}

Durante a execução deste estudo, um conjunto de trabalhos foi publicados na
forma de artigos e notas técnicas. Este apêndice traz uma descrição deste
material em ordem cronológica reversa.

\paragraph{Agosto de 2006: \sectioneng{Deployment of the ATLAS High-Level
Trigger}}, IEEE Transactions on Nuclear Science. Referência~\ref{aa:tns-06}

\begin{quotation}
The ATLAS combined test beam in the second half of 2004 saw the first
deployment of the ATLAS High-Level Trigger (HLT). The next steps are
deployment on the pre-series farms in the experimental area during 2005,
commissioning and cosmics tests with the full detector in 2006 and collisions
in 2007. This paper reviews the experience gained in the test beam, describes
the current status and discusses the further enhancements to be made. We
address issues related to the dataflow, integration of selection algorithms,
testing, software distribution, installation and improvements.
\end{quotation}

\paragraph{Maio de 2006: \sectioneng{A configuration system for the ATLAS
trigger}}, Journal of Instrumentation, Maio de
2006. Referência~\ref{aa:jinst-06}

\begin{quotation}
The ATLAS detector at CERN's Large Hadron Collider will be exposed to
proton-proton collisions from beams crossing at 40 MHz that have to be
reduced to the few hundreds of Hz allowed by the storage systems. A
three-level trigger system has been designed to achieve this goal. We describe
the configuration system under construction for the ATLAS trigger chain. It
provides the trigger system with all the parameters required for decision
taking and to record its history. The same system configures the event
reconstruction, Monte Carlo simulation and data analysis, and provides tools
for accessing and manipulating the configuration data in all contexts.
\end{quotation}

\paragraph{Abril de 2006: \sectioneng{Neural triggering system operating on
high resolution calorimetry information}} na Nuclear Instruments and Methods
in Physics Research. Referência~\cite{aa:nim-06}

\begin{quotation}
This paper presents an electron/jet discriminator system for operating at the
Second Level Trigger of ATLAS. The system processes calorimetry data and
organizes the regions of interest in the calorimeter in the form of concentric
ring sums of energy deposition, so that both signal compaction and high
performance can be achieved. The ring information is fed into a feed forward
neural discriminator. This implementation resulted on a 97\% electron
detection efficiency for a false alarm of 3\%. The full discrimination chain
could still be executed in less than 500 us.
\end{quotation}

\paragraph{Fevereiro de 2006: \setioneng{Testing on a Large Scale : running
the ATLAS Data Acquisition and High Level Trigger Software on 700 PC Nodes}}
Computing In High Energy and Nuclear Physics  CHEP 2006 , Mumbai, India , 13 -
17 Feb 2006. Referência~\cite{aa:chep-06-01}

\begin{quotation}
The ATLAS Data Acquisition (DAQ) and High Level Trigger (HLT) software system
will be comprised initially of 2000 PC nodes which take part in the control,
event readout, second level trigger and event filter operations. This high
number of PCs will only be purchased before data taking in 2007. The large
CERN IT LXBATCH facility provided the opportunity to run in July 2005 online
functionality tests over a period of 5 weeks on a stepwise increasing farm
size from 100 up to 700 PC dual nodes. The interplay between the control and
monitoring software with the event readout, event building and the trigger
software has been exercised the first time as an integrated system on this
large scale. New was also to run algorithms in the online environment for the
trigger selection and in the event filter processing tasks on a larger
scale. A mechanism has been developed to package the offline software together
with the DAQ/HLT software and to distribute it via peer-to-peer software
efficiently to this large pc cluster. The findings obtained during the tests
lead to many immediate improvements in the software. Trend analysis allowed
identifying critical areas. Running an online system on a cluster of 700 nodes
successfully was found to be especially sensitive to the reliability of the
farm as well as the DAQ/HLT system itself and the future development will
concentrate on fault tolerance and stability.
\end{quotation}

\paragraph{Julho de 2005: \sectioneng{Overview of the High-Level Trigger
Electron and Photon Selection for the ATLAS Experiment at the LHC}} na 14th
IEEE - NPSS Real Time Conference 2005 Nuclear Plasma Sciences
Society. Referência~\cite{aa:rt-05-01}

\begin{quotation}
The ATLAS experiment at the Large Hadron Collider (LHC) will face the
challenge of efficiently selecting interesting candidate events in pp
collisions at 14 TeV center-of-mass energy, whilst rejecting the enormous
number of background events. The High-Level Trigger (HLT = second level
trigger and Event Filter), which is a software based trigger will need to
reduce the level-1 output rate of ~75 kHz to ~200 Hz written out to mass
storage. In this talk an overview of the current physics and system
performance of the HLT selection for electrons and photons is given. The
performance has been evaluated using Monte Carlo simulations and has been
partly demonstrated in the ATLAS testbeam in 2004. The efficiency for the
signal channels, the rate expected for the selection, the global data
preparation and execution times will be highlighted. Furthermore, some physics
examples will be discussed to demonstrate that the triggers are well adapted
for the physics programme envisaged at the LHC.
\end{quotation}

\paragraph{Julho de 2005: \sectioneng{Configuration of the ATLAS trigger}} na
14th IEEE - NPSS Real Time Conference 2005 Nuclear Plasma Sciences
Society. Referência~\cite{aa:rt-05}

\begin{quotation}
The ATLAS detector at CERN's LHC will be exposed to proton-proton collisions
at a rate of 40 MHz. In order to reduce the data rate to about 200 Hz, only
potentially interesting events are selected by a three-level trigger
system. Its first level is implemented in electronics and firmware whereas the
higher trigger levels are based on software. To prepare the full trigger chain
for the online event selection according to a certain strategy, a system is
being set up that provides the relevant configuration information -
e.g. values for hardware registers in level-1 or parameters of high-level
trigger algorithms - and stores the corresponding history. The same
information is used to configure the offline trigger simulation. In this
presentation an overview of the ATLAS trigger system is given concentrating on
the event selection strategy and its description. The technical implementation
of the configuration system is summarized.
\end{quotation}

\paragraph{Setembro de 2004: \sectioneng{Implementation and Performance of the
High Level Trigger Electron and Photon Selection for the ATLAS Experiment at
the LHC}}, na conferência Computing in High Energy Physics and Nuclear
Physics, em Interlaken, na Suíça. Referência~\cite{aa:chep-04-01}

\begin{quotation}
The ATLAS experiment at the Large Hadron Collider (LHC) will face the
challenge of efficiently selecting interesting candidate events in pp
collisions at 14 TeV center of mass energy, while rejecting the enormous
number of background events, stemming from an interaction rate of up to 10^9
Hz. The First Level trigger will reduce this rate to around O(100
kHz). Subsequently, the High Level Trigger (HLT), which is comprised of the
Second Level trigger and the Event Filter, will need to further reduce this
rate by a factor of O(10^3). The HLT selection is software based and will be
implemented on commercial CPUs, using a common framework built on the standard
ATLAS object oriented software architecture. In this paper an overview of the
current implementation of the selection for electrons and photons in the HLT
is given. The performance of this implementation has been evaluated using
Monte Carlo simulations in terms of the efficiency for the signal channels,
rate expected for the selection, data preparation times, and algorithm
execution times. Besides the efficiency and rate estimates, some physics
examples will be discussed, showing that the triggers are well adapted for the
physics programme envisaged at LHC. The electron and photon trigger software
is also being exercised at the ATLAS 2004 Combined Test Beam, where components
from all ATLAS subdetectors are taking data together along the H8 SPS
extraction line; from these tests a validation of the selection architecture
chosen in a real on-line environment is expected.
\end{quotation}

\paragraph{Junho de 2004: \sectioneng{Algorithms for the ATLAS high-level
trigger}}. Referência~\cite{aa:tns-04-05}.

\begin{quotation}
Following rigorous software design and analysis methods, an object-based
architecture has been developed to derive the second- and third-level trigger
decisions for the future ATLAS detector at the LHC. The functional components
within this system responsible for generating elements of the trigger
decisions are algorithms running within the software architecture. Relevant
aspects of the architecture are reviewed along with concrete examples of
specific algorithms and their performance in "vertical" slices of various
physics selection strategies.
\end{quotation}

\paragraph{Junho de 2004: \sectioneng{The base-line DataFlow system of the ATLAS
Trigger and DAQ}} Na revista \sectioneng{IEEE Transactions on Nuclear
Science}, volume 51, número 3. Referência \cite{aa:tns-2004-3}.

\paragraph{Junho de 2004: \sectioneng{ATLAS TDAQ data collection software} na
\sectioneng{IEEE Transactions on Nuclear Science}, volume 51, número
3. \cite{aa:tns-2004-3}.

\begin{quotation}
The DataCollection (DC) is a subsystem of the ATLAS Trigger and DAQ system. It
is responsible for the movement of event data from the ReadOut subsystem to
the Second Level Trigger and to the Event Filter. This functionality is
distributed on several software applications running on Linux PCs
interconnected with Gigabit Ethernet. For the design and implementation of
these applications a common approach has been adopted. This approach leads to
the design and implementation of a common DC software framework providing a
suite of common services. 10 Refs.
\end{quotation}

\paragraph{Outubro de 2003: \sectioneng{Architecture of the ATLAS Online
Physics-Selection Software at the LHC}.} Este artigo foi apresentado no
\sectioneng{8th ICATPP Conference Proceedings} em Villa Erba, Como, Itália. 
Referência \cite{aa:como-2003}.

@@

\paragraph{Março de 2003: \sectioneng{Architecture of the ATLAS High Level
Trigger Event SelectionSoftware}} Este trabalho foi apresentado no congresso
\sectioneng{Computing for High-Energy Physics' 2003}, em \sectioneng{La
Jolla}, Califórnia, EUA. Referência \cite{aa:chep-2003-3}

@@

\paragraph{Agosto de 1998: \sectioneng{A neural online triggering system based
on parallel processing}} Publicado na IEEE Transactions on Nuclear Science.

\begin{quotation}
The study of a prototype of the second-level triggering system for operation
at LHC conditions is addressed by means of a parallel machine
implementation. The 16 node transputer based machine uses a fast digital
signal processor acting as a coprocessor for optimizing signal processing
applications. A C-language development environment is used for running all
applications at ultimate speed. The implementation is based on information
supplied by four detectors and includes two phases of system operation:
feature extraction and global decision. Feature extraction for calorimeters
and global decision processing are performed by means of neural
networks. Preprocessing and neural network parameters rest in memory and the
activation function is implemented using a look up table. Simulated data for
the second-level trigger operation are used for performance evaluation.
\end{quotation}
